---
title: "Assignment1"
author: "Jiaqi Sun"
format: html
editor: visual
---

## Problem 1

(a) Import the data and name the columns.

```{r}
wine <- read.table("D:/Code/R/wine/wine.data", sep = ",", header = FALSE)

colnames(wine) <- c("Class", "Alcohol", "Malic_acid", "Ash", "Alcalinity_of_ash", 
                    "Magnesium", "Total_phenols", "Flavanoids", 
                    "Nonflavanoid_phenols", "Proanthocyanins", 
                    "Color_intensity", "Hue", "OD280/OD315_of_diluted_wines", 
                    "Proline")
```

(b) **Class1: 59, Class2: 71, Class3: 48**. Numbers are correct.

```{r}
table(wine$Class)
```

\(c\) 1. The wine with the highest alcohol content belongs to **Class1**.

```{r}
max_alcohol <- max(wine$Alcohol)
class_max_alcohol <- wine[wine$Alcohol ==  max_alcohol, 1]
print(paste("Highest alcohol is in class:", class_max_alcohol))
```

\(c\) 2. The wine with the lowest alcohol content belongs to **Class2**.

```{r}
min_alcohol <- min(wine$Alcohol)
class_min_alcohol <- wine[wine$Alcohol ==  min_alcohol, 1]
print(paste("Lowest alcohol is in class:", class_min_alcohol))
```

\(c\) 3. There are **26 wines** that have higher levels of magnesium.

```{r}
table(wine$Magnesium > 114)
```

\(c\) 4. The number of wines that have higher magnesium in each class are **Class1: 15, Class2: 6, Class3: 5.**

```{r}
table(wine[wine$Magnesium > 114, 1])
```

\(d\) The first row is the average value of each variable, the second to forth rows are average value of each variable within each class.

```{r}
overall_ave <- colMeans(wine[, -1])
class_1_ave <- colMeans(wine[wine$Class == 1, -1])
class_2_ave <- colMeans(wine[wine$Class == 2, -1])
class_3_ave <- colMeans(wine[wine$Class == 3, -1])
average_tabel <- rbind(overall_ave, class_1_ave, class_2_ave, class_3_ave)
average_tabel
```

\(e\) **My own calculation is down below.** Here are the built-in function carrying out the t-test.

```{r}
t.test(wine[wine$Class == 1, "Ash"], wine[wine$Class == 2, "Ash"])
t.test(wine[wine$Class == 1, "Ash"], wine[wine$Class == 3, "Ash"])
t.test(wine[wine$Class == 2, "Ash"], wine[wine$Class == 3, "Ash"])
```

My own calculation. Note that the t-test is:

$$
t = \frac{\bar X_1-\bar X_2}{\sqrt{\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2}}} \\
df  \approx \frac{(\frac{s_1^2}{n_1}+\frac{s_2^2}{n_2})^2}{\frac{s_1^4}{n_1^2(n_1-1)}+\frac{s_2^4}{n_2^2(n_2-1)}}
$$

```{r}
# my own code for t test
class_number <- table(wine$Class)
n <- vector(length = 3)
sample_mean <- vector(length = 3)
sample_var <- vector(length = 3)
wine_Ash <- wine[, c(1, 4)]

for(i in 1:3){
  n[i] <- as.integer((class_number[i])) # get the number for each class
  sample_mean[i] <- mean(wine[wine$Class == i, "Ash"]) # calculate mean
  sample_var[i] <- var(wine[wine$Class == i, "Ash"]) # calculate variance
}
# t statistic
t12 <- (sample_mean[1] - sample_mean[2]) / sqrt(sample_var[1]/n[1] + sample_var[2]/n[2])
t13 <- (sample_mean[1] - sample_mean[3]) / sqrt(sample_var[1]/n[1] + sample_var[3]/n[3])
t23 <- (sample_mean[2] - sample_mean[3]) / sqrt(sample_var[2]/n[2] + sample_var[3]/n[3])

# degree of freedom
df12 <- (sample_var[1]/n[1]+sample_var[2]/n[2])^2/(sample_var[1]^2/n[1]^2/(n[1]-1)+sample_var[2]^2/n[2]^2/(n[2]-1))
df13 <- (sample_var[1]/n[1]+sample_var[3]/n[3])^2/(sample_var[1]^2/n[1]^2/(n[1]-1)+sample_var[3]^2/n[3]^2/(n[3]-1))
df23 <- (sample_var[2]/n[2]+sample_var[3]/n[3])^2/(sample_var[2]^2/n[2]^2/(n[2]-1)+sample_var[3]^2/n[3]^2/(n[3]-1))

# calculate the p-value
p12 <- (1 - pt(abs(t12), df = df12)) * 2
p13 <- (1 - pt(abs(t13), df = df13)) * 2
p23 <- (1 - pt(abs(t23), df = df23)) * 2
cat("The t statistic between class 1 and class 2 is:", t12, "with p-value:", p12)
cat("The t statistic between class 1 and class 3 is:", t13, "with p-value:", p13)
cat("The t statistic between class 2 and class 3 is:", t23, "with p-value:", p23)
```

Interpretation:

The t-statistic between Class 1 and Class 2 is 4.418367, and the p-value is 2.124341e-05, which is less than 0.05. There is strong evidence to reject the null hypothesis. Class 1 and Class 2 are not at the same level of Ash.

The t-statistic between Class 1 and Class 3 is 0.4648899, and the p-value is 0.642973, which is greater than 0.05. There is no evidence to reject the null hypothesis. Class 1 and Class 3 are at the same level of Ash.

The t-statistic between Class 2 and Class 3 is -4.183958, and the p-value is 5.626653e-05, which is less than 0.05. There is strong evidence to reject the null hypothesis. Class 2 and Class 3 are not at the same level of Ash.

## Problem 2

\(a\)

```{r}
#' is perfect power function
#'
#' @param x a positive integer we are going to check
#' @param power a positive integer, the power value
#'
#' @return a list of length 2. The first entry is a logical value indicating whether the number is a perfect power. The second entry is the root. If the number is not perfect power, root = -1
#'
#' @examples > isPerfectPower(27, power = 3)
#'           $isPerfect
#'           [1] TRUE
#'           $root
#'           [1] 3
#'
isPerfectPower <- function(x, power) {
  # make sure that number and power are positive
  if (x <= 0 || power <= 0) {
    cat("number and power should be positive integer")
    return(list(isPerfect = FALSE, root = -1))
  }
  
  root <- 2
  while (root^power <= x) {
    # loop through all natural number with the power, until the value is larger than the number.
    if (root^power == x) {
      return(list(isPerfect = TRUE, root = root))
    }
    root <- root + 1
  }
  # no root is found, then return FALSE and root = -1.
  return(list(FALSE, -1))
}

isPerfectPower(4, 2)
isPerfectPower(1000, 2)
isPerfectPower(5066499666, 4979)
isPerfectPower(1024, 10)
isPerfectPower(2048, 11)
```

\(b\)

```{r}
#' find root power function
#'
#' @param x a positive integer we are going to find root
#'
#' @return if number is perfect power, return the string equation. If not, return NULL
#'
#' @examples > findRootPower(125)
#'           [1] "125 = 5^3"
#'
findRootPower <- function(x) {
  # make sure that number is greater than 1.
  if (x <= 1) {
    cat("number should be greater than 1.\n")
    return(invisible(NULL))
  }
  
  power <- 2
  while (2^power <= x) {
    # loop through all the powers, until the smallest perfect power is larger than the number.
    result <- isPerfectPower(x, power)
    if (isTRUE(result$isPerfect)) {
      return(paste(x, " = ", result$root, "^", power, sep = ""))
    }
    power <- power + 1
  }
  # no root is found
  cat(x, "is not a perfect power.\n")
  return(invisible(NULL))
}
```

```{r}
findRootPower(27)
findRootPower(13060694016)
findRootPower(7776)
findRootPower(170859375)
findRootPower(58247422)
findRootPower(94143178827)
```

## Problem 3

> To accomplish this task, we'll create two R functions. The first function will determine the name of a 5-card hand in 5-card stud poker, and the second function will simulate dealing a round of cards for a specified number of players and display the hands along with their names. Here are the two functions:
>
> ```{r}
> # Function to determine the name of a 5-card poker hand
> get_poker_hand_name <- function(suits, ranks) {
>   # Ensure that the input vectors are of length 5
>   if (length(suits) != 5 || length(ranks) != 5) {
>     stop("Both suits and ranks vectors should have a length of 5.")
>   }
>   
>   # Count the frequency of each rank
>   rank_counts <- table(ranks)
>   
>   # Check for flush and straight
>   is_flush <- length(unique(suits)) == 1
>   is_straight <- length(rank_counts) == 5 && (max(ranks) - min(ranks) == 4)
>   
>   # Determine the hand name
>   if (is_flush && is_straight) {
>     return("Straight Flush")
>   } else if (max(rank_counts) == 4) {
>     return("Four of a Kind")
>   } else if (max(rank_counts) == 3 && length(rank_counts) == 2) {
>     return("Full House")
>   } else if (is_flush) {
>     return("Flush")
>   } else if (is_straight) {
>     return("Straight")
>   } else if (max(rank_counts) == 3) {
>     return("Three of a Kind")
>   } else if (length(rank_counts) == 3) {
>     return("Two Pair")
>   } else if (length(rank_counts) == 4) {
>     return("One Pair")
>   } else {
>     return("High Card")
>   }
> }
> >
> # Function to simulate a round of 5-card stud poker
> simulate_poker_round <- function(num_players) {
>   # Define the suits and ranks
>   suits <- c("Hearts", "Diamonds", "Clubs", "Spades")
>   ranks <- c(2:10, "Jack", "Queen", "King", "Ace")
>   
>   # Create a deck of cards
>   deck <- expand.grid(Rank = ranks, Suit = suits)
>   
>   # Shuffle the deck
>   deck <- deck[sample(nrow(deck)), ]
>   
>   # Deal 5 cards to each player
>   players <- list()
>   for (i in 1:num_players) {
>     hand <- deck[((i - 1) * 5 + 1):(i * 5), ]
>     players[[i]] <- hand
>     hand_name <- get_poker_hand_name(hand$Suit, hand$Rank)
>     cat("Player", i, "Hand:", paste(hand$Rank, hand$Suit), "-", hand_name, "\n")
>   }
> }
> >
> # Example usage:
> simulate_poker_round(4)  # Simulate a round with 4 players
> >
> ```
>
> In this code, the \`get_poker_hand_name\` function takes in vectors of suits and ranks, checks the hand for various poker hand combinations, and returns the name of the hand. The \`simulate_poker_round\` function simulates a round of 5-card stud poker for a specified number of players, deals hands to each player, and displays the hands along with their names using the first function.

-   Are the inputs and outputs as described above?

-   Are the hands valid (e.g. real cards, no duplicates, right number)?

-   Are the names of the hands correct? (See [here](https://www.cardplayer.com/rules-of-poker/hand-rankings) if you're not familiar with poker hands.)

-   Does it ensure no duplicates in cards across hands? What happens if you ask for more than 10 hands to be dealt (as there are only 52 cards in a standard deck)?
